---
title: "Food Inspections Report"
author: Andy Choens
date: April 28, 2017
output: 
    html_document:
        toc: true
        depth: 1
---

# About

```{r init, echo=FALSE, message=FALSE, warning=FALSE}

## PACKAGES ----------------------------
## Adding the message=FALSE and warning=FALSE prevents dplyr's
## start-up messages do not become part of the final report.
library(dplyr, lib="~/lib")
library(pander, lib="~/lib")
library(rmarkdown, lib="~/lib")

## DATA --------------------------------
load("data/inspections.Rda")

## LOCAL VARS --------------------------
data_date <- as.character( unique(inspections$data_date), format="%m/%d/%Y")

```

This is a simple demonstrates the power and flexibility of RMarkdown
to produce integrated, reproducible reports. The code to produce all
of the tables, charts, etc. in this document are embedded in the
food-inspection-report.Rmd file.

New York State, and other states, regularly performs inspections of
establishments serving food to the public. The data to run this report
was downloaded from
the
[NYS Open Data](https://dev.socrata.com/foundry/health.data.ny.gov/f285-x9ha) site
on `r data_date`. On that date, 
`r format(nrow(inspections), big.mark=",")` 
inspections were downloaded.

This report intends to answer two question: 

- Which county has the highest number of violations, and 
- Which county has the highest violation rate?

# Violations

For the purposes of this report, the number of violations reported by
county is the total of critical and non-critical food-safety
violations reported. The violation rate is calculated as the
percentage of food inspections which resulted in at least one
food-safety violation.

## Most Violations

The following table shows, by county, the total number of food
inspections performed in the current data release and the total number
of violations found. Below, the five counties from the current
food-inspections data-set with the highest total number of violations.

```{r violations, echo=FALSE}

viol_tot_county <-
    inspections %>%
    group_by(county) %>%
    summarize(n_insp          = n(),
              tot_violations  = sum(total_violations)
              ) %>%
    arrange(desc(tot_violations)) %>%
    select("County" = county,
           "N Inspections"    = n_insp,
           "Total Violations" = tot_violations)

## This adds a comma to note the thousands.
viol_tot_county[,"Total Violations"] <- format(viol_tot_county[,"Total Violations"], big.mark = ",")

pander(viol_tot_county[1:5,],
       caption = "Top five counties with the highest number of total violations",
       justify="lrr")

```

## Violation Rate

The following table shows, by county, the total number of food
inspections performed in the current data release, and the violation
rate. The violation rate is defined as the percentage of
food-inspections which find at least one violation. Below, the five
counties with the highest rate of violations. It should not suprise
you if these are very similar to the previous table.

```{r violation-rate, echo=FALSE}

viol_rate_county <-
    inspections %>%
    group_by(county) %>%
    summarize(n_insp          = n(),
              rate_violations = round(100*sum(total_violations > 1) / n(),2)
              )  %>%
    arrange(desc(rate_violations)) %>%
    select("County"         = county,
           "N Inspections"    = n_insp,
           "Violation Rate" = rate_violations)

pander(viol_rate_county[1:5,],
       caption = "Top five counties with the highest rate of violations",
       justify="lrr")

```
